\texttt{davos}: The Python package smuggler

### 2022-04-06 paper/main.tex
finish motivation

Python has become one of the most widely used and fastest-growing scientific programming languages by combining an accessible, high-level syntax with the availability of powerful third-party tools that facilitate rapid development and collaboration \cite{MullEtal15}. The Python ecosystem offers an extensive data science toolkit, with platforms for interactive programming (e.g., Project Jupyter \cite{KluyEtal16}, Google Colaboratory), community-maintained libraries for data manipulation (e.g., NumPy \cite{HarrEtal20}, SciPy \cite{VirtEtal20}, Pandas \cite{McKi10}) and visualization (e.g., Matplotlib \cite{Hunt07}, seaborn \cite{Wask21}), frameworks for training complex machine learning models (e.g., scikit-learn \cite{PedrEtal11}, TensorFlow \cite{AbadEtal15}, Hugging Face \cite{WolfEtal20}), and myriad other resources. However, this heavy emphasis on third-party libraries also presents a challenge to writing and sharing stable, reproducible scientific Python code: different versions of the same package may behave differently, use different syntax, add or drop support for specific functions or other libraries, address (or introduce) bugs, and so on. While these issues are present to some extent in any language or ecosystem, they have a particular impact on the Python community due to its unusually rapid growth relative to other languages. Ensuring Python code behaves consistently over time and across users therefore typically requires making sure it is always run with a specific set of versions for each package used.

One common approach to solving this problem is to create containerized or virtualized Python environments (e.g., using \href{https://www.docker.com/}{Docker}, \href{https://sylabs.io/singularity/}{Singularity}, or \href{https://docs.conda.io/en/latest/}{conda}) tailored to individual applications. Researchers can then share these environments alongside their code as configuration files that explicitly list required packages and versions, allowing other users to build identical copies for themselves. While effective, this approach bears two significant drawbacks: First it can substantially add to the technical knowledge, system resources, and setup required to share and run the encapsulated code.

One common approach to solving this problem is to create containerized or virtualized Python environments (e.g., using \href{https://www.docker.com/}{Docker}, \href{https://sylabs.io/singularity/}{Singularity}, or \href{https://docs.conda.io/en/latest/}{conda}) tailored to individual applications. Researchers can then share these environments alongside their code as configuration files that explicitly list required packages and package versions, allowing others to build identical copies themselves. While powerful, this approach bears two significant drawbacks: First, it adds to the

One common approach to solving this problem involves creating containerized or virtualized Python environments (e.g., using \href{https://www.docker.com/}{Docker}, \href{https://sylabs.io/singularity/}{Singularity}, or \href{https://docs.conda.io/en/latest/}{conda}) tailored to individual applications. Authors may then share these environments alongside their code as configuration files from which users may build identical copies themselves. While effective, a significant drawback to this approach is that it introduces additional prerequisite knowledge

% While effective, such tools can add a significant degree of complexity to the process of sharing research code for both the author and

a significant drawback to this approach is that it adds to the requisite knowledge

in many cases, it introduces a level of complexity beyond . For example, distributing research code that relies on a particular Docker image to run properly not only necessitates extra configuration files and setup steps, but requires that both the author and end user install and navigate additional software that is often more complicated and resource-intensive than the actual code being shared. These added prerequisites clash with the simplicity and accessibility that have helped popularize Python among researchers, and can create barriers to both contributing to and taking advantage of open science. \texttt{davos} was developed with the goal of addressing

>>>>

Python has become one of the most widely used and fastest-growing scientific programming languages by combining an accessible, high-level syntax with the availability of powerful third-party tools that facilitate rapid development and collaboration \cite{MullEtal15}. The Python ecosystem offers an extensive data science toolkit, with platforms for interactive programming (e.g., Project Jupyter \cite{KluyEtal16}, Google Colaboratory), community-maintained libraries for data manipulation (e.g., NumPy \cite{HarrEtal20}, SciPy \cite{VirtEtal20}, Pandas \cite{McKi10}) and visualization (e.g., Matplotlib \cite{Hunt07}, seaborn \cite{Wask21}), frameworks for training complex machine learning models (e.g., scikit-learn \cite{PedrEtal11}, TensorFlow \cite{AbadEtal15}, Hugging Face \cite{WolfEtal20}), and myriad other resources. However, this heavy emphasis on third-party libraries also presents a challenge to writing and sharing stable, reproducible scientific Python code: different versions of the same library may behave differently, adopting different syntax, adding or dropping support for specific functions or other libraries, addressing (or introducing) bugs, and so on. While these issues are present to some extent in any language or ecosystem, they have a particular impact on the Python community due to its unusually rapid growth relative to other languages. Ensuring Python code behaves consistently over time and across users therefore typically requires making sure it is always run with the same specific set of versions for each package used.

One common approach to solving this problem is to create containerized or virtualized Python environments (e.g., using \href{https://www.docker.com/}{Docker}, \href{https://sylabs.io/singularity/}{Singularity}, or \href{https://docs.conda.io/en/latest/}{conda}) tailored to individual applications. Researchers can then share these environments alongside their code as configuration files that explicitly list required packages and versions, enabling other users to build identical copies for themselves. However, this approach comes with two major drawbacks: First, it can add significantly to the technical knowledge, system resources, and initial setup required to share and run the actual code of interest. For example, sharing research code that relies on a particular Docker image to run properly not only necessitates distributing extra configuration files and setup instructions, but also requires that both the original author and anyone with whom the code may be shared install and navigate additional software that is likely more complex and resource-intensive than the Python code it supports. These added prerequisites clash with the simplicity and accessibility that have contributed to Python's popularity among scientists, and can create significant barriers to both contributing to and taking advantage of open science resources.

Second, typical environment management tools don't enforce that a particular requirements \textit{remain} satisfied after they are initially installed. While this provides a degree of flexibility in certain use cases, it makes it easy to inadvertently alter the set of installed Python packages after this initial setup. For example, suppose a researcher has implemented a series of analyses using version 1.0 of ``Package \textit{X}," and decides to  perform an additional analysis requiring ``Package \textit{Y}." If Package \textit{Y} depends on version 1.1 of Package \textit{X}, then Package \textit{X} will be upgraded to accommodate this new requirement, potentially leading to inconsistent behavior of code that uses Package \textit{X} if this change goes unnoticed.
